Relocating a Cluster of Earthquakes Using a Single Seismic Station

### 2012-12-17 paper2/paper2.tex
more changes

We design a 2D synthetic acoustic experiment (example 1) by randomly
selecting $\hat{x}$- and $\hat{y}$-coordinates such that $-50 \leq
\hat{x}, \hat{y} \leq 50$\,m. These are indicated with triangles in
Figure \ref{fig-2D50eq-relocation_eg1}.
 We assume a local velocity of $\alpha=3,300\,$ms$^{-1}$ between all
event pairs and a dominant frequency of 2.5$\,$Hz to represent
waveform data filtered between 1 and 5$\,$Hz. The CWI data are
defined by the dominant wavelength normalized positive bounded
Gaussian PDF with statistics $\bar{\mu}_N$ and $\bar{\sigma}_N$
\citep{dr_Robinson11a}. A hypothetical CWI mean is created by
setting
\begin{equation}
\label{eq:hypothetical-CWI-mean-optichapt}
\bar{\mu}_N = \mu_1 \left( \widetilde{\delta}_t \right)
\end{equation}
using equation (\ref{eq:mu1}). This assumption ensures that the
sample mean of hypothetical separation estimates is consistent with
known CWI biases \citep{dr_Robinson11a}. In example 1 we use
$\bar{\sigma}_N=0.02$ between all event pairs. Application of our
optimization procedure on the hypothetical CWI data yields the
circles in Figure \ref{fig-2D50eq-relocation_eg1}. The optimization
does not lead to the exact solution due to the addition of noise
($\bar{\sigma}_N=0.02$) on the hypothetical CWI data. The average
coordinate error is 2.0\,m which is small compared to the noise of
$\bar{\sigma}_N=0.02$ which for $v_s=3300$\,ms$^{-1}$ and
$f_{dom}=2.5$\,Hz corresponds to roughly 25\,m.
\begin{figure}
\includegraphics[width = 20pc]{diags/locs_2D_50eq_1.eps}
\caption{Example 1 - Synthetic relocation of 50 earthquakes in 2D
using all constraints with noise $\bar{\sigma}_N=0.02$. Actual and
optimization event locations are shown in triangles and circles,
respectively.} \label{fig-2D50eq-relocation_eg1}
\end{figure}

\begin{figure}
\includegraphics[width = 20pc]{diags/locs_2D_50eq_3.eps}
\caption{Example 2 - Synthetic relocation of 50 earthquakes in 2D
using all constraints with noise $\bar{\sigma}_N= 2
\epsilon(\delta_t)$.
 Actual and optimization event locations
are shown in triangles and circles, respectively.}
\label{fig-2D50eq-relocation_eg3}
\end{figure}
\citet{dr_Robinson11a} demonstrates that the noise on CWI estimates
is often larger than 0.02 and that it increases with event
separation. Consequently, example 1 is simplistic because we fix
$\bar{\sigma}_N=0.02$ for all pairs. In example 2 we increase the
uncertainty and introduce a distance dependance into the
hypothetical $\bar{\sigma}_N$ by defining
$\bar{\sigma}_N=\epsilon(\delta_t)$, where
$\epsilon(\delta_t)$ is the half-width of the errorbars for a
synthetic acoustic experiment with filtering between 1 and 5\,Hz
\citep[see Fig. 4(b) of ][]{dr_Robinson11a}. Repeating the
optimization leads to the circles in Figure
\ref{fig-2D50eq-relocation_eg3} which have an average coordinate
error of 2.8\,m.

@@

\begin{figure}
%\noindent\includegraphics[width = 20pc]{../../thesis_version2/diags/eq_location_optimisation/synth2Dmulti/ressummary_2Dsynth50eq.eps}
\noindent\includegraphics[width =
20pc]{diags/synth2Dmulti/ressummary_2Dsynth50eq.eps}
\caption{Example 3 - Statistical measures of error in the
 solutions for the 2D synthetic cases when all and best
optimization results are considered. The statistics $\Delta_{max}$
and $\Delta_\mu$  are the maximum and mean coordinate error,
respectively. The bottom subplot shows the average minimum number of
branches required to link the 2450 pairs.}
 \label{fig:optimisationresults-2Dsynth}
\end{figure}

@@

\begin{figure}
%\noindent\includegraphics[width = 20pc]{../../thesis_version2/diags/eq_location_optimisation/synth3Dmulti/ressummary_3Dsynth50eq.eps}
\noindent\includegraphics[width =
20pc]{diags/synth3Dmulti/ressummary_3Dsynth50eq.eps}
\caption{Example 4 - Statistical measures of error in the
optimization solutions for the 3D synthetic cases when all (gray)
and best (black) results are considered. The statistics
$\Delta_{max}$ and $\Delta_\mu$ are the maximum and mean coordinate
error, respectively. The absence of the gray and black lines below
60\% and 30\% indicates a breakdown in the solutions when all or
best optimization result(s) are considered, respectively.}
\label{fig:optimisationresults-3Dsynth}
\end{figure}

@@

\begin{figure}
\noindent\includegraphics[width =
20pc]{diags/CalaverasMap/gmt_california/CaliforniaCalaverasMap1.eps}
\caption{Map showing location of the Calaveras
cluster (white star) and 805 seismic stations (black triangles).}
\label{fig:-eqopti-California-Calaveras}
\end{figure}

@@

%%=======================================================================================
% This is the original figure using HypoDD with LSQR
%\begin{figure*}
%\includegraphics{diags/CalaverasLoc1.eps}
%\caption{Comparison of earthquake hypocenters using three different methods: catalogue location (column 1), hypoDD with all
%308 events (column 2) and CWI example 1 (column 3).
%Note that in the case of the CWI locations we consider only the 68 earthquakes in red, the
%blue events are shown for the purpose of orientation only.}
%\label{fig-69Calaverasevents_eg1}
%\end{figure*}

@@

\begin{figure}
\noindent\includegraphics[width =
20pc]{diags/CalaverasMap/matlab/Calaveras_substationmap}
\caption{Location of the 10 stations (triangles) used to relocate
the Calaveras events in Example 6 to 8. Stations are removed one at
a time according to the order in Table
\ref{tab:Calaveras-stationremoval} and the events relocated. The
events are indicated with black circles.}
\label{fig:-eqopti-Calaveras-substations}
\end{figure}

@@

\begin{figure*}
\includegraphics[angle=90,height = 50pc]{diags/CalaverasLoc2.eps}
\caption{Example 6 - CWI relative locations with reduced stations.}
\label{fig-CWIreducesstats}
\end{figure*}

@@

%%=========================================================================
% this is the original version using hypoDD with LSQ
%\begin{figure*}
%\includegraphics[angle=90,height = 50pc]{diags/CalaverasLoc3.eps}
%\caption{....}
%\label{fig-HYPODDreducesstats}
%\end{figure*}

@@

%%=========================================================================
% This is the original version using hypoDD with LSQ
%\begin{figure}
%\noindent\includegraphics[width = 20pc]{diags/CalaverasLoc4.eps}
%\caption{Statistics on coordinate differences for reduced station inversions. Differences
%are computed between the inversion results (CWI and hypoDD) and the complete hypoDD
%locations for all 308 events. The top subplot illustrates the number of constrainable
%events in the CWI and hypoDD inversions as a function of the stations considered.}
%\label{fig-statremoval_summarystats}
%\end{figure}

% This is the version using hypoDD with SVD
\begin{figure}
\noindent\includegraphics[width =
20pc]{diags/CalaverasLoc4_hypoDD_SVD.eps} \caption{Example 6 -
Number of constrainable events $nE$ in the CWI and hypoDD inversions
as a function of the stations considered (top). Mean (middle) and
maximum (bottom) of the difference computed between the reduced
station inversion results (CWI and hypoDD) and the complete hypoDD
locations for all 308 events. } \label{fig-statremoval_summarystats}
\end{figure}
%%=========================================================================

@@

%%=========================================================================
% This is the original version using hypoDD with LSQ
% This is only makes sense for LSQR so we remove it from this manuscript
%\begin{figure*}
%\noindent\includegraphics{diags/CalaverasLoc5.eps}
%\caption{Locations of the 68 Calaveras earthquakes combining all available coda wave and travel time constraints with three different
%levels of uncertainty on the travel time PDFs. Column 1 sets $\sigma_x = \sigma_y = 19.5$\,m and $\sigma_z = 15$\,m
%\citep[after]{dr_Waldhauser08a}, Column 2 uses $\sigma_x = \sigma_y = 3 \times19.5$\,m and $\sigma_z = 3\times15$\,m and
%column 3 considers $\sigma_x = \sigma_y = 152$\,m and $\sigma_z = 232$\,m \citep[after]{dr_Shearer97a}. }
%\label{fig-68Calaverasevents_ttandcoda1}
%\end{figure*}

@@

\begin{figure}
\noindent\includegraphics[width = 20pc]{diags/OmoriFigure.eps}
\caption{Cumulative number of aftershocks for the
Hokkaido-Nansei-Oki, Japan $M_s=7.8$ earthquake of 12 July 1993
using equation (\ref{eq:CumOmori}). The leftmost gray, middle black
and rightmost gray lines signify aftershocks occurring before,
during and after the deployment of a pseudo temporary array
installed 4 days after the main shock and left for 150 days. A
temporary deployment of this kind will record roughly 50\% of the
aftershocks in the 1000 days following the mainshock. }
\label{fig:Omorifigure}
\end{figure}

@@

%%=========================================================================
% This is the original version using hypoDD with LSQ
%\begin{figure}
%\noindent\includegraphics{diags/CalaverasLoc6.eps}
%\caption{Left - Combined travel time and coda wave inversions using travel time constraints on
%22 (or 32\%) of the events and coda waves from station CCO only. All travel time based uncertainties
%are assigned $\sigma_x = \sigma_y = 3 \times19.5$\,m and $\sigma_z = 3\times15$\,m. Right - travel time
%relocation only (i.e. no CWI) using ........}
%\label{fig-68Calaverasevents_ttsubsetandcoda1}
%\end{figure}

@@

\appendix

>>>>

We design a 2D synthetic acoustic experiment (example 1) by randomly
selecting $\hat{x}$- and $\hat{y}$-coordinates such that $-50 \leq
\hat{x}, \hat{y} \leq 50$\,m. These are indicated with triangles in
Figure \ref{fig-2D50eq-relocation_eg1}.
 We assume a local velocity of $\alpha=3,300\,$ms$^{-1}$ between all
event pairs and a dominant frequency of 2.5$\,$Hz to represent
waveform data filtered between 1 and 5$\,$Hz. The CWI data are
defined by the dominant wavelength normalized positive bounded
Gaussian PDF with statistics $\bar{\mu}_N$ and $\bar{\sigma}_N$
\citep{dr_Robinson11a}. A hypothetical CWI mean is created by
setting
\begin{equation}
\label{eq:hypothetical-CWI-mean-optichapt}
\bar{\mu}_N = \mu_1 \left( \widetilde{\delta}_t \right)
\end{equation}
using equation (\ref{eq:mu1}). This assumption ensures that the
sample mean of hypothetical separation estimates is consistent with
known CWI biases \citep{dr_Robinson11a}. In example 1 we use
$\bar{\sigma}_N=0.02$ between all event pairs. Application of our
optimization procedure on the hypothetical CWI data yields the
circles in Figure \ref{fig-2D50eq-relocation_eg1}. The optimization
does not lead to the exact solution due to the addition of noise
($\bar{\sigma}_N=0.02$) on the hypothetical CWI data. The average
coordinate error is 2.0\,m which is small compared to the noise of
$\bar{\sigma}_N=0.02$ which for $v_s=3300$\,ms$^{-1}$ and
$f_{dom}=2.5$\,Hz corresponds to roughly 25\,m.

\citet{dr_Robinson11a} demonstrates that the noise on CWI estimates
is often larger than 0.02 and that it increases with event
separation. Consequently, example 1 is simplistic because we fix
$\bar{\sigma}_N=0.02$ for all pairs. In example 2 we increase the
uncertainty and introduce a distance dependance into the
hypothetical $\bar{\sigma}_N$ by defining
$\bar{\sigma}_N=\epsilon(\delta_t)$, where
$\epsilon(\delta_t)$ is the half-width of the errorbars for a
synthetic acoustic experiment with filtering between 1 and 5\,Hz
\citep[see Fig. 4(b) of ][]{dr_Robinson11a}. Repeating the
optimization leads to the circles in Figure
\ref{fig-2D50eq-relocation_eg3} which have an average coordinate
error of 2.8\,m.

@@

@@

@@

@@

@@

@@

@@

@@

@@

@@

@@

@@

%%=======================================================================================
% Table 1

%%=======================================================================================
% Table 2

%%=======================================================================================
% Table 3

%%=======================================================================================

%Figure 1
\begin{figure}
%\includegraphics[width = 20pc]{diags/locs_2D_50eq_1.eps}
\includegraphics[width = 20pc]{Figure1_c.eps}
\caption{Example 1 - Synthetic relocation of 50 earthquakes in 2D
using all constraints with noise $\bar{\sigma}_N=0.02$. Actual and
optimization event locations are shown in triangles and circles,
respectively.} \label{fig-2D50eq-relocation_eg1}
\end{figure}

%%=======================================================================================

%Figure 2
\begin{figure}
%\includegraphics[width = 20pc]{diags/locs_2D_50eq_3.eps}
\includegraphics[width = 20pc]{Figure2_c.eps}
\caption{Example 2 - Synthetic relocation of 50 earthquakes in 2D
using all constraints with noise $\bar{\sigma}_N= 2
\epsilon(\delta_t)$.
 Actual and optimization event locations
are shown in triangles and circles, respectively.}
\label{fig-2D50eq-relocation_eg3}
\end{figure}

%%=======================================================================================

% Figure 3
\begin{figure}
%\noindent\includegraphics[width = 20pc]{../../thesis_version2/diags/eq_location_optimisation/synth2Dmulti/ressummary_2Dsynth50eq.eps}
%\noindent\includegraphics[width=20pc]{diags/synth2Dmulti/ressummary_2Dsynth50eq.eps}
\noindent\includegraphics[width=20pc]{Figure3_c.eps}
\caption{Example 3 - Statistical measures of error in the
 solutions for the 2D synthetic cases when all and best
optimization results are considered. The statistics $\Delta_{max}$
and $\Delta_\mu$  are the maximum and mean coordinate error,
respectively. The bottom subplot shows the average minimum number of
branches required to link the 2450 pairs.}
 \label{fig:optimisationresults-2Dsynth}
\end{figure}

%%=======================================================================================

% Figure 4
\begin{figure}
%\noindent\includegraphics[width = 20pc]{../../thesis_version2/diags/eq_location_optimisation/synth3Dmulti/ressummary_3Dsynth50eq.eps}
%\noindent\includegraphics[width=20pc]{diags/synth3Dmulti/ressummary_3Dsynth50eq.eps}
\noindent\includegraphics[width=20pc]{Figure4_c.eps}
\caption{Example 4 - Statistical measures of error in the
optimization solutions for the 3D synthetic cases when all (gray)
and best (black) results are considered. The statistics
$\Delta_{max}$ and $\Delta_\mu$ are the maximum and mean coordinate
error, respectively. The absence of the gray and black lines below
60\% and 30\% indicates a breakdown in the solutions when all or
best optimization result(s) are considered, respectively.}
\label{fig:optimisationresults-3Dsynth}
\end{figure}

%%=======================================================================================

% Figure 5
\begin{figure}
\noindent\includegraphics[width =
20pc]{diags/CalaverasMap/gmt_california/CaliforniaCalaverasMap1.eps}
\caption{Map showing location of the Calaveras cluster (white star)
and 805 seismic stations (black triangles).}
\label{fig:-eqopti-California-Calaveras}
\end{figure}

%%=======================================================================================
% Figure 6

% This is the original figure using HypoDD with LSQR
%\begin{figure*}
%\includegraphics{diags/CalaverasLoc1.eps}
%\caption{Comparison of earthquake hypocenters using three different methods: catalogue location (column 1), hypoDD with all
%308 events (column 2) and CWI example 1 (column 3).
%Note that in the case of the CWI locations we consider only the 68 earthquakes in red, the
%blue events are shown for the purpose of orientation only.}
%\label{fig-69Calaverasevents_eg1}
%\end{figure*}

% Figure 7
\begin{figure}
%\noindent\includegraphics[width=20pc]{diags/CalaverasMap/matlab/Calaveras_substationmap}
\noindent\includegraphics[width=20pc]{Figure7.eps} \caption{Location
of the 10 stations (triangles) used to relocate the Calaveras events
in Example 6 to 8. Stations are removed one at a time according to
the order in Table \ref{tab:Calaveras-stationremoval} and the events
relocated. The events are indicated with black circles.}
\label{fig:-eqopti-Calaveras-substations}
\end{figure}

%%=======================================================================================

%Figure 8
\begin{figure*}
\includegraphics[angle=90,height = 50pc]{diags/CalaverasLoc2.eps}
\caption{Example 6 - CWI relative locations with reduced stations.}
\label{fig-CWIreducesstats}
\end{figure*}

%%=========================================================================
% Figure 9

% this is the original version using hypoDD with LSQ
%\begin{figure*}
%\includegraphics[angle=90,height = 50pc]{diags/CalaverasLoc3.eps}
%\caption{....}
%\label{fig-HYPODDreducesstats}
%\end{figure*}

%Figure 10

% This is the original version using hypoDD with LSQ
%\begin{figure}
%\noindent\includegraphics[width = 20pc]{diags/CalaverasLoc4.eps}
%\caption{Statistics on coordinate differences for reduced station inversions. Differences
%are computed between the inversion results (CWI and hypoDD) and the complete hypoDD
%locations for all 308 events. The top subplot illustrates the number of constrainable
%events in the CWI and hypoDD inversions as a function of the stations considered.}
%\label{fig-statremoval_summarystats}
%\end{figure}

% This is the version using hypoDD with SVD
\begin{figure}
\noindent\includegraphics[width =
20pc]{diags/CalaverasLoc4_hypoDD_SVD.eps} \caption{Example 6 -
Number of constrainable events $nE$ in the CWI and hypoDD inversions
as a function of the stations considered (top). Mean (middle) and
maximum (bottom) of the difference computed between the reduced
station inversion results (CWI and hypoDD) and the complete hypoDD
locations for all 308 events. } \label{fig-statremoval_summarystats}
\end{figure}

% Figure 11

% This is the original version using hypoDD with LSQ
% This is only makes sense for LSQR so we remove it from this manuscript
%\begin{figure*}
%\noindent\includegraphics{diags/CalaverasLoc5.eps}
%\caption{Locations of the 68 Calaveras earthquakes combining all available coda wave and travel time constraints with three different
%levels of uncertainty on the travel time PDFs. Column 1 sets $\sigma_x = \sigma_y = 19.5$\,m and $\sigma_z = 15$\,m
%\citep[after]{dr_Waldhauser08a}, Column 2 uses $\sigma_x = \sigma_y = 3 \times19.5$\,m and $\sigma_z = 3\times15$\,m and
%column 3 considers $\sigma_x = \sigma_y = 152$\,m and $\sigma_z = 232$\,m \citep[after]{dr_Shearer97a}. }
%\label{fig-68Calaverasevents_ttandcoda1}
%\end{figure*}

% Figure 12

\begin{figure}
%\noindent\includegraphics[width = 20pc]{diags/OmoriFigure.eps}
\noindent\includegraphics[width = 20pc]{Figure12_c.eps}
\caption{Cumulative number of aftershocks for the
Hokkaido-Nansei-Oki, Japan $M_s=7.8$ earthquake of 12 July 1993
using equation (\ref{eq:CumOmori}). The leftmost gray, middle black
and rightmost gray lines signify aftershocks occurring before,
during and after the deployment of a pseudo temporary array
installed 4 days after the main shock and left for 150 days. A
temporary deployment of this kind will record roughly 50\% of the
aftershocks in the 1000 days following the mainshock. }
\label{fig:Omorifigure}
\end{figure}

% Figure 13

% This is the original version using hypoDD with LSQ
%\begin{figure}
%\noindent\includegraphics{diags/CalaverasLoc6.eps}
%\caption{Left - Combined travel time and coda wave inversions using travel time constraints on
%22 (or 32\%) of the events and coda waves from station CCO only. All travel time based uncertainties
%are assigned $\sigma_x = \sigma_y = 3 \times19.5$\,m and $\sigma_z = 3\times15$\,m. Right - travel time
%relocation only (i.e. no CWI) using ........}
%\label{fig-68Calaverasevents_ttsubsetandcoda1}
%\end{figure}

\appendix
\clearpage
